---
title: "Spatial Pattern Analysis of Siphonophore Vertical Habitat Use"
author: "Alejandro DamianSerrano"
date: "3/8/2017"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sp)
library(spatial)
library(spatstat)
library(reshape2)
library(rmutil)
library(scatterplot3d)
library(vegan)
library(fields)
library(gstat)
library(statnet)
library(car)
library(stpp)
library(metafor)
```

```{r dataload, include=F}
zdata = read.csv("~/Dropbox/siphonophore_phylogeny_2017/character_coding/Siphonophore_depth_pruned.tsv", sep='\t')
zdata[,-1] = lapply(zdata[,-1],as.numeric)
rownames(zdata) = zdata$Species

funnel(zdata$Median-mean(zdata$Median), sei=zdata$StdDev/sqrt(zdata$Count))
funnel(zdata$Median-mean(zdata$Median), vi=1000000/sqrt(zdata$Count))
```

##Goal:

Simulate and compare 3D spatial distribution patterns of different siphonophore species and other zooplankton (potential prey) from ROV annotations' bathymetric summary statistics and patchiness metrics.

Calculate distance distributions between different planktonic taxa.

Calculate relative encounter probabilities for each pair of siphonophore-prey species.

Estimate an expected dietary covariance matrix for siphonophore species pairs given their distributions relative to prey

## Siphonophore Bathymetric Profile Similarity

Loaded ROV depth data including `{r} colnames(zdata)`
Transform variables to numeric. Make rownames = Species names.

```{r explore, echo=F}
#map(zdata[,-1], hist)
heatmap(as.matrix(zdata[,c(-1,-5)]))
```

##Simulated Normal Distribution

Generate normal distributions of vertical depth data from the real mean and standard deviation, with an N equivalent to the real abundance counts. Produced a list of simulated observations of different species in different depths.

```{r simulate, include=T, warning=FALSE, message=FALSE}
znorm  = list()
for(i in 1:nrow(zdata)){
  print(zdata$Species[i])
  #print(rnorm(zdata$Count[i], zdata$Mean[i], zdata$StdDev[i]))
  zi = rnorm(zdata$Count[i], zdata$Mean[i], zdata$StdDev[i])
  znorm[[i]] = zi[which(zi>zdata$Min[i], zi<zdata$Max[i])]
}
znormX=rep(0,max(zdata$Count))
for(i in 1:length(znorm)){
  zni = c(znorm[[i]],rep(NA, (max(zdata$Count) - length(znorm[i]))))
  znormX = rbind(znormX,zni)
}
dim(znormX)
znorm_data = melt(cbind(zdata$Species,data.frame(znormX[-1,1:250])))
znorm_data = znorm_data[,-2]
names(znorm_data) = c("Species","Depth")
```

Barplot and boxplot from ggplot2 help visualize the data simulated.

```{r plots, echo=T}
znorm_data$Depth = -1*znorm_data$Depth
p <- ggplot(znorm_data, aes(y=Depth,x=Species)) + geom_boxplot()
p + theme(axis.text.x = element_text(angle = 90, hjust = 1))
p2 <- ggplot(znorm_data, aes(x=Depth,fill=Species),xlim=1) + geom_histogram(binwidth = 20, alpha=.4)
p2
```

##Spatial XYZ Simulations

*How do we estimate predator-prey encounter rates?*

Encounter rate: E = B(encounter rate kernel) C1(predator density) C2(prey density)
B = 3/2 pi L(Predator colony length) (Prey length)

Next chunk:
  Random generation of X-Y values in an arbitrary reference system in meters, ascribed to each Z-Species observation. Product: List of dataframes for each species with XYZ data.
  Plotted 3D ocean space with clouds of distributed siphonophore species.
  By Species Distance List: List of dataframes for each species with Euclidean distances between observations of the same species.

```{r spatsim3d, fig.height=18, fig.width=18, echo=T}

X = runif(nrow(znorm_data), max=1000)
Y = runif(nrow(znorm_data), max=1000)
xyzdata = cbind(znorm_data, X, Y)[,c(1,3,4,2)]
xyzdata$Depth = xyzdata$Depth*-1
s3d <- with(xyzdata, scatterplot3d(X, Y, Depth, color = as.numeric(Species), pch = 19))
legend(s3d$xyz.convert(0.5, 0.7, 0.5), pch = 19, yjust=0, legend = levels(xyzdata$Species), col = colors()[400:429])
split(x=xyzdata[which(!is.na(xyzdata$Depth)),], f=xyzdata[which(!is.na(xyzdata$Depth)),]$Species) -> byspp_xyz
lapply(byspp_xyz, function(x) { x["Species"] <- NULL; x }) %>% lapply(vegdist, method="euclidean") %>% lapply(as.matrix) -> byspp_dists
xyz_noNA = xyzdata[which(!is.na(xyzdata$Depth)),]
```

Plotted X(artbitrary horizontal component)-Z(depth) plots with combinations of species.

```{r spatsim2d, fig.height=18, fig.width=18, echo=F}
xzdata = xyz_noNA[,c(1,2,4)]

plot(xzdata[which(xzdata$Species == 'Nanomia bijuga'),-1], col='red', cex=3, pch='N', ylim=c(2000,0))
points(xzdata[which(xzdata$Species == 'Praya dubia'),-1], col='blue', , cex=3, pch='P')
points(xzdata[which(xzdata$Species == 'Marrus claudanielis'),-1], col='orange', cex=3, pch='M')
points(xzdata[which(xzdata$Species == 'Kephyes ovata'),-1], col='green', , cex=3, pch='K')

#lapply(byspp_dists, heatmap) -> heatmapsX
#vegdist(xyzdata[which(!is.na(xyzdata$Depth)),2:4], method="euclidean") %>% as.matrix %>% heatmap
```

##Spatial Analysis in 2D

Generate PPP objects using sp package.
Calculate spatial sumary statistics: Spherical contact

Take a 500 points sample of the large XYZ matrix, calculate euclidean pairwise distances between species (from the set of points that belong to each species using a unique identifier loop), plot heatmap.

```{r spat_analysis2D, fig.height=18, fig.width=18, echo=T}
xz_S = SpatialPointsDataFrame(coords=xzdata[-1],data=xzdata[1], match.ID=T)
as.ppp(xzdata[-1], c(-5000,5000,-5000,5000)) %>% Hest() -> sphcont
sphcont$hazard

xyz_noNA  -> xyzSample #%>% sample_n(1000)
xyzSample[-1] %>% vegdist(method="euclidean") %>% as.matrix -> xyzDistSample
rownames(xyzDistSample) = colnames(xyzDistSample) = xyzSample$Species

SpeciesPairWise <- function(dists){
  Species = unique(rownames(dists))
  Iterations = 1:length(unique(rownames(dists)))
  PairwiseDistances = matrix(nrow=length(Iterations), ncol = length(Iterations))
  for(i in Iterations){
   sp1=Species[i]
   for(j in Iterations){
     sp2=Species[j]
     PairwiseDistances[i,j] = mean(dists[which(rownames(dists) == sp1),which(colnames(dists) == sp2)])
  }}
  return(as.matrix(PairwiseDistances))
  }

xyzPWSample = SpeciesPairWise(xyzDistSample) #Pairwise distances between species, 3D sample
rownames(xyzPWSample) = colnames(xyzPWSample) = unique(rownames(xyzDistSample))
PairwisePxyz = 1 - xyzPWSample/max(xyzPWSample)
#PairwisePxyz %>% heatmap()

xzSample = xyzSample[,-3] #same as above but with a 2D sample
xzSample[-1] %>% vegdist(method="euclidean") %>% as.matrix -> xzDistSample
rownames(xzDistSample) = colnames(xzDistSample) = xzSample$Species
xzPWSample = SpeciesPairWise(xzDistSample)
rownames(xzPWSample) = colnames(xzPWSample) = unique(rownames(xzDistSample))
PairwisePxz = 1 - xzPWSample/max(xzPWSample)

zSample = xyzSample[,c(-2,-3)] #same as above but with a 1D sample (only Depth)
zSample[-1] %>% vegdist(method="euclidean") %>% as.matrix -> zDistSample
rownames(zDistSample) = colnames(zDistSample) = zSample$Species
zPWSample = SpeciesPairWise(zDistSample)
rownames(zPWSample) = colnames(zPWSample) = unique(rownames(zDistSample))
PairwisePz = 1 - zPWSample/max(zPWSample)

#Lets test whether adding random X and Y dimensions adds any information
mantel(PairwisePxyz, PairwisePxz)
mantel(PairwisePxz, PairwisePz)
mantel(PairwisePxyz, PairwisePz)

#and the winner is... they all have the exact same signal!
#So, we stick with the simplest solution: zDist

melt(zDistSample) -> zDistSampleMelt
colnames(zDistSampleMelt) = c("SP1", "SP2", "Dist")

zDistSampleMelt[which(zDistSampleMelt$Dist < 0.001 & zDistSampleMelt$Dist > 0 & zDistSampleMelt$SP1 != zDistSampleMelt$SP2), ] # %>% .[which((.$SP1 %in% Predators & .$SP2 %in% Prey)|(.$SP1 %in% Prey & .$SP2 %in% Predators)), ]
```
##Pairwise Correlation function (Rippley's K)
```{r K_analysis2D, fig.height=18, fig.width=18, echo=T}
xzSample[1:2000,2:3] %>% as.ppp(owin(c(min(xzSample[2]),max(xzSample[2])), c(min(xzSample[3]),max(xzSample[3])))) -> xzPPP
xzPPP %>% pcf() -> xzPCF
xzPPP %>% Kest() -> xzKest
#plot(envelope(xzPPP,pcf))
#plot(envelope(xzPPP,Kest))
marks(xzPPP) <- xzSample[1:2000,1]
plot(Smooth(xzPPP))
cut.ppp(xzPPP) -> xzMPPP
Gdot(xzPPP, i="Nanomia bijuga") %>% plot()
Kcross(xzPPP, i="Nanomia bijuga", j="Praya dubia", r=c(0,1000))
KcrossData = c("SP1", "SP2", "ISO", "TRANS")
for(i in xzPPP$marks[1:30]){
  print(i)
  for(j in xzPPP$marks[1:30]){
    if(i!=j){
      A = Kcross(xzPPP, i,j, r=c(0,1000))
      class(A) = "data.frame"
      B = c(i,j,mean(A$iso),mean(A$trans))
      KcrossData = rbind(KcrossData,B)
    }
  }
}
colnames(KcrossData) = KcrossData[1,]
KcrossData = as.data.frame(KcrossData[-1,])
rownames(KcrossData) = 1:nrow(KcrossData)
KcrossData = mutate(KcrossData, paste(KcrossData$SP1,KcrossData$SP2))
names(KcrossData)[5] = "SPP"
plot = ggplot(KcrossData) + geom_point(aes(x=ISO, y=TRANS, color=SPP)) + scale_color_discrete(guide = FALSE)

split.ppp(xzPPP) -> split_xzPPP
split_xzPPP$`Apolemia lanosa`$y %>% summary()
```


##Simulate patchy distributions
```{r cluster, fig.height=18, fig.width=18, echo=F}
xyzSampleList = split(xyzSample, xyzSample$Species)
xyzSampleClust = xyzSampleList
for(i in 1:length(xyzSampleList)){
  rMatClust(.01, runif(1,3,20), 100, win=owin(c(0,100),c(0,100))) %>% as.data.frame() -> XYi
  xyzSampleClust[[i]][1:nrow(xyzSampleList[[i]]),c(2,3)] = XYi
}
xyzSampleClust <- do.call(rbind, xyzSampleClust)
rownames(xyzSampleClust) = 1:nrow(xyzSampleClust)
#ggplot(xyzSampleClust[1:1000,], aes(x=X,y=-Depth,color=Species)) + geom_point()
s3d <- with(xyzSampleClust, scatterplot3d(X, Y, -Depth, color = as.numeric(Species), pch = 19))
legend(s3d$xyz.convert(0.5, 0.7, 0.5), pch = 19, yjust=0, legend = levels(xyzSampleClust$Species), col = colors()[400:429])

xyzSampleClust[,-1] %>% vegdist(method="euclidean") %>% as.matrix -> xyzClustDist
rownames(xyzClustDist) = colnames(xyzClustDist) = xyzSampleClust$Species

xyzPWClust = SpeciesPairWise(xyzClustDist) #Pairwise distances between species, 3D sample
rownames(xyzPWClust) = colnames(xyzPWClust) = unique(rownames(xyzClustDist))
PairwisePClustxyz = 1 - xyzPWClust/max(xyzPWClust)

mantel(PairwisePxyz, PairwisePClustxyz)
mantel(PairwisePClustxyz, PairwisePz)

```

##Depth Bins Analysis
```{r bins}
xyzSplit100 = split(xyzSample, cut(xyzSample$Depth, seq(0, max(xyzSample$Depth),100)))  
names(xyzSplit100) = paste(seq(0, max(xyzSample$Depth)-100, 100), seq(100, max(xyzSample$Depth), 100), sep=" - ")

xyzDistSplit100 = list()
ind = 0
for(E in xyzSplit100){
  ind = ind+1
  bin = as.matrix(dist(E[,-1]))
  rownames(bin) = colnames(bin) = E$Species
  xyzDistSplit100[[ind]] = SpeciesPairWise(bin)
  rownames(xyzDistSplit100[[ind]]) = colnames(xyzDistSplit100[[ind]]) = unique(rownames(bin))
  print(xyzDistSplit100[[ind]])
  }

```


**To do list:**
-Obtain prey distribution profiles
-Obtain raw Z depth data for siphonophores
-Obtain patchiness metrics from ROV transect data
-Calculate other spatial summary statistics per pair of species
  Rippley's K, Pairwise Correlation Function




